---
title: "P8106_HW1"
author: "Ziyi Zhao"
date: "2/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(summarytools)
library(stargazer)
library(leaps)
library(caret)
library(FNN)
library(ModelMetrics)
library(Rcpp)
library(microbenchmark)

```


# Part a: Fit a linear model using least squares on the training data and calculate the mean square error using the test data.

We first look at the summary statistics of predictors and the response.

```{r}
sol_train = read_csv("./solubility_train.csv") %>% janitor::clean_names()
sol_test = read_csv("./solubility_test.csv") %>% janitor::clean_names()

st_options(plain.ascii = FALSE,
           style = "rmarkdown",
           dfSummary.silent = TRUE,
           footnote = NA,
           subtitle.emphasis = FALSE)

dfSummary(sol_train[,-229])
dfSummary(sol_test[,-229])

sol_all <- rbind(sol_train,sol_test)
dfSummary(sol_all[,-229])

```

We fit linear least square model to train dataset and calculate MSE using test dataset.

```{r}
linear_fit <- lm(solubility~.,data=sol_train)

linear_pred <- predict(linear_fit,sol_test)

mse(sol_test$solubility,linear_pred)

# KNN
mse_knn <- rep(0,19)
for (i in 2:20) {
  pred_knn <- knn.reg(train = sol_train[,1:228], test = sol_test[,1:228],
                     y = sol_train$solubility,k=i)
  mse_knn[i-1] <- mse(sol_test$solubility,pred_knn$pred)
}

min(mse_knn)

knn_k <- c(2:20)
tibble(knn_k,mse_knn) %>% 
  ggplot(aes(x=knn_k,y=mse_knn))+
  geom_point()+
  labs(title = "MSE change by different k",
       x = "k",
       y = "mse")


```


# Part b: Fit a ridge regression model on the training data, with λ chosen by cross-validation. Report the test error

```{r}

```







